---
title: "Homework 2"
author: "Jeanette Shekelle"
date: "10/5/2018"
output: github_document
---
```{r}
library(tidyverse)
```

## Problem 1


Import NYC transit data. Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable

```{r}
NYC_transit_data = 
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select (line, station_name, station_latitude, station_longitude, route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11, entry, vending, ada) %>%
mutate (entry = recode(entry, "YES"= TRUE, "NO"= FALSE))

```


Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

The dataset contains 19 variables

How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St A/B/C/D; 125st 1; 125st 4/5);
```{r}
nrow(distinct(NYC_transit_data, station_name, line))
```

There are 465 dinstinct stations. 


How many stations are ADA compliant?

```{r}
nrow(filter(distinct(NYC_transit_data, station_name, line, ada), ada == TRUE))
```

There are 84 ADA compliant stations among distinct stations. 


What proportion of station entrances / exits without vending allow entrance?

```{r}
nrow(filter(NYC_transit_data, vending == "NO", entry == TRUE)) / nrow(filter(NYC_transit_data, vending == "NO"))
```

The proportion is 0.377. 


Reformat data so that route number and route name are distinct variables. 

```{r}
NYC_transit_tidy_data = gather(NYC_transit_data, key = route_number, value = route_name, route1:route11)
NYC_transit_tidy_data
```


How many distinct stations serve the A train? 

```{r}
nrow(filter(distinct(NYC_transit_tidy_data, station_name, route_name, line), route_name == "A"))
```

60 distinct stations serve the A train. 

Of the stations that serve the A train, how many are ADA compliant?

```{r}
nrow(filter(filter(distinct(NYC_transit_tidy_data, station_name, route_name, line, ada), route_name == "A"), ada == TRUE))
```

Only 17 stations are ADA compliant.



## Problem 2


Read and clean the Mr. Trash Wheel sheet:

- specify the sheet in the Excel file and to omit columns containing notes (using the range argument and cell_cols() function)
- use reasonable variable names
- omit rows that do not include dumpster-specific data
- rounds the number of sports balls to the nearest integer and converts the result to an integer variable (using  as.integer)


```{r}
library(readxl)
trash_wheel_data = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", "Mr. Trash Wheel", range = "A2:N338")  %>% 
janitor::clean_names() %>% 
    filter(!is.na(dumpster)) %>% 
mutate(sports_balls = round(sports_balls)) %>% 
  mutate(sports_balss = as.integer(sports_balls))
```


Read and clean precipitation data for 2016 and 2017. For each, omit rows without precipitation data and add a variable year. Next, combine datasets and convert month to a character variable (the variable  month.name is built into R and should be useful).

```{r}
precip_16 = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", "2016 Precipitation", range = "A2:B14")  %>% 
 janitor::clean_names() %>% 
  mutate(year = 2016)

precip_17 = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", "2017 Precipitation", range = "A2:B14")  %>% 
 janitor::clean_names() %>% 
  mutate(year = 2017)

precip_16_17= bind_rows(mutate(precip_16, month = month.name), mutate(precip_17, month = month.name))

```

The number of observations in the trash wheel dataset is `r nrow(trash_wheel_data)`. The number of variables is `r ncol(trash_wheel_data)`. In 2017, the total precipitation was `r sum(filter(precip_16_17, year == 2017)$total)`. In 2016, the total precipitation was `r sum(filter(precip_16_17, year == 2016)$total)`. The median number of sports balls in a dumpster in 2016 was `r median(filter(trash_wheel_data, year == 2016)$sports_balls)`. 

## Problem 3

```{r}
# install.packages("devtools")
devtools::install_github("p8105/p8105.datasets")
```

```{r}
library(p8105.datasets)
```

```{r}
data("brfss_smart2010")
```

For this question:

- format the data to use appropriate variable names;
- focus on the “Overall Health” topic
- exclude variables for class, topic, question, sample size, and everything from lower confidence limit to GeoLocation
- structure data so that values for Response (“Excellent” to “Poor”) are column names / variables which indicate the proportion of subjects with each response (which are values of Data_value in the original dataset)
- create a new variable showing the proportion of responses that were “Excellent” or “Very Good”

```{r}
brfss_smart2010 = 
  brfss_smart2010 %>% 
   janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  select(- class, - topic, - question, - sample_size, - (confidence_limit_low:geo_location)) %>% 
  spread(., key = response, value = data_value) %>% 
  janitor::clean_names() %>% 
  mutate(excellent_very_good = excellent + very_good)
 

```


How many unique locations are included in the dataset? Is every state represented? What state is observed the most?

There are `r length(unique(brfss_smart2010$locationdesc))` unique locations included in the dataset. Every state is represented; there are `r length(unique(brfss_smart2010$locationabbr))` unique values in the state abbreviation column. The state of `r tail(names(sort(table(brfss_smart2010$locationabbr))), 1)` is observed the most. 


- In 2002, what is the median of the “Excellent” response value?

```{r}
brfss_smart2010 %>% 
  filter(year == 2002) %>% 
  summarize(median(excellent, na.rm = TRUE))
```

The median of the excellent response values is 23.6. 

- Make a histogram of “Excellent” response values in the year 2002.

```{r}
brfss_smart2010 %>% 
  filter(year == 2002) %>% 
ggplot(aes(x = excellent)) + 
  geom_histogram(binwidth = 0.6, fill = "Pink") +
  labs(title = "Excellent Response Values in 2002", x = "Excellent Response", y = "Frequency", caption = "Data: BRFSS Smart 2012") +
  theme_gray()
```


- Make a scatterplot showing the proportion of “Excellent” response values in New York County and Queens County (both in NY State) in each year from 2002 to 2010.

```{r}
brfss_smart2010 %>% 
  filter(locationdesc == "NY - New York County" | locationdesc == "NY - Queens County") %>% 
  ggplot(aes(x = year, y = excellent)) +
geom_point(aes(color = locationdesc), alpha = .5) +
labs(title = "Excellent Response Values in 2002: New York County and Queens County", x = "Year", y = "Proportion with Excellent Response", caption = "Data: BRFSS Smart 2012") +
  theme_bw()

```

