---
title: "Homework 2"
author: "Jeanette Shekelle"
date: "9/30/2018"
output: github_document
---
```{r}
library(tidyverse)
```

## Problem 1


Import NYC transit data. Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable

```{r}
NYC_transit_data = 
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names()%>%
  select (line, station_name, station_latitude, station_longitude, route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11, entry, vending, ada) %>%
  
mutate (
 recode(entry, "YES"= TRUE, "NO"= FALSE)
  ) 

```


Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

The dataset contains 19 variables

How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St A/B/C/D; 125st 1; 125st 4/5);
```{r}
nrow(distinct(NYC_transit_data, station_name, line))
```

There are 465 dinstinct stations. 


How many stations are ADA compliant?

```{r}
count(NYC_transit_data, ada == TRUE)
```

There are 468 ADA compliant stations. 


What proportion of station entrances / exits without vending allow entrance?

```{r}
nrow(
  NYC_transit_data %>%
filter(entry == TRUE) %>%
filter(vending == "NO") 
/
nrow( NYC_transit_data %>%
    filter(vending == "NO")
))

```


## Problem 2


Read and clean the Mr. Trash Wheel sheet:

- specify the sheet in the Excel file and to omit columns containing notes (using the range argument and cell_cols() function)
- use reasonable variable names
- omit rows that do not include dumpster-specific data
- rounds the number of sports balls to the nearest integer and converts the result to an integer variable (using  as.integer)

```{r}
library(readxl)
trash_wheel_data = read_excel("data/HealthyHarborWaterWheelTotals2018-7-28.xlsx")
res <- read.xlsx(file, 1) 
head(trash_wheel_data, 5)

```



